Computation and comparison of the stable Northeastern US marine boundary layer

### 2020-12-01 Paper/methodology.tex
Add governing equations section

@@

\subsubsection{Governing equations}
\fxnote{Shreyas write this part.  Currently placeholder equations below.}\\

The low-Mach number equations:
\begin{align}
  \frac{\partial \rho} {\partial t} + \frac{\partial \rho u_j}{\partial x_j} & = 0, \\
\frac{\partial \rho u_i}{\partial t} + \frac{\partial \rho u_j u_i}{\partial x_j}
+ \frac{\partial P}{\partial x_i} & = \frac{\partial \tau_{ij}}{\partial x_j}
+ \left( \rho - \rho_{\circ} \right) g_i, \\
\frac{\partial \rho h}{\partial t} + \frac{\partial \rho u_j h}{\partial x_j} & =
- \frac{\partial q_j}{\partial x_j} + \frac{\partial P_{th}}{\partial t},
\end{align}

@@

\subsubsection{Nalu-Wind}
\fxnote{Shreyas write this part}\\
Describe Nalu-Wind

>>>>

Large Eddy Simulations (LES) of the stable atmopsheric conditions described in
Table~\ref{tab:CapeWindMeasurements} are performed using the open-source,
ExaWind simulation environment\footnote{\url{https://github.com/exawind}}.
ExaWind is a suite of high-fidelity modeling tools for analyzing complex
microscale atmospheric boundary layer flows and wind farm flows. The present
study uses two different incompressible computational fluid dynamic solvers
available within the ExaWind suite --
Nalu-Wind~\footnote{\url{https://github.com/exawind/nalu-wind}} and
AMR-Wind~\footnote{\url{https://github.com/exawind/amr-wind}}. This section
briefly describes the incompressible Navier-Stokes equations solved in the
solvers, the specifics of the turbulence model and the wall shear stress
formulation used to model the terrain, and the details of the numerical
discretization of the two equations in the two codes.

@@

\subsubsection{Governing equations}
Both CFD codes solve the incompressible form of Navier-Stokes equations with
appropriate models for turbulence closure. Equation~\ref{eqn:ns-les} shows the
continuity and momentum equations with all the terms necessary to model the
atmospheric boundary layer problem. Term $\mathbf{I}$ represents the pressure
gradient forces as a deviation from hydrostatic and horizontal mean gradient,
term $\mathbf{II}$ represents the contribution from viscous and sub-filter scale
stresses, term $\mathbf{III}$ represents the contribution from Coriolis forces
due to earth's rotation, term $\mathbf{IV}$ represents the effects of buoyancy,
and term $\mathbf{V}$ represents source terms necessary to drive the flow to a
desired horizontal mean velocity at prescribed heights.

\begin{align}
  \frac{\partial \bar{\rho}} {\partial t} + \frac{\partial \bar{\rho} \widetilde{u_j}}{\partial x_j} & = 0, \nonumber\\
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{u}_i\right) +
  \frac{\partial}{\partial x_j} \left( \bar{\rho}\, \widetilde{u}_i \widetilde{u}_j \right) &=
  - \underbrace{\frac{\partial p'}{\partial x_j} \delta_{ij}}_\mathbf{I}
  - \underbrace{\frac{\partial \tau_{ij}}{\partial x_j}}_\mathbf{II}
  - \underbrace{2\bar{\rho}\,\epsilon_{ijk}\,\Omega_j\widetilde{u_k}}_\mathbf{III}
  + \underbrace{\left(\rho - \rho_\circ \right) g_i}_\mathbf{IV}
  + \underbrace{S^{u}_{i}}_\mathbf{V} \label{eqn:ns-les}
\end{align}

For the simulations presented in this paper, the buoyancy term ($\mathbf{IV}$)
is approximated using the Bousinessq model. The Bousinessq model ignores the
effect of density gradients on inertia while retaining its effects on buoyancy.
The density fluctuations governing the buoyancy effects are approximated as

\begin{align*}
  \frac{\rho}{\rho_\circ} \approx 1 - \beta \left( \theta - \theta_\circ \right)
\end{align*}
This leads to a forcing term that depends on potential temperature ($\theta$),
the reference density $\rho_\circ = \bar{\rho}$, and thermal expansion
coefficient $\beta \approx 1 / \theta_\circ$ as
\begin{align}
  \left(\rho - \rho_\circ \right) g_i = -\rho_\circ\, \beta\, g_i \left( \theta - \theta_\circ \right)
\end{align}

For ABL problems, the conservation of energy equation is usually written in
terms of potential temperature as shown in Eq.~\ref{eqn:pot-temp-les}.
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{\theta}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{\theta} \right) = - \frac{\partial}{\partial x_j} \hat{q}_j \label{eqn:pot-temp-les}
\end{align}
Under the assumption of ideal gas conditions and constant specific heat capacity
$c_p$, the gradients in potential temperature are proportional to the gradients
in absolute temperature, i.e.,
\begin{align*}
   \left[ \frac{\partial T}{\partial t}, \frac{\partial T}{\partial x}, \frac{\partial T}{\partial y} \right] =
   \left( \frac{\bar{p}}{p_\circ} \right)^{\left(\frac{R}{c_p}\right)} \left[ \frac{\partial \theta}{\partial t}, \frac{\partial \theta}{\partial x}, \frac{\partial \theta}{\partial y} \right]
\end{align*}
Furthermore, ignoring the pressure and viscous work terms in
Eq.~\ref{eqn:pot-temp-les}, for incompressible flows, it can be shown that
solving the enthalpy equation is equivalent to solving the potential temperature
equation. The enthalpy equation solved in Nalu-Wind for wind energy problems is
shown below
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{h}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{h} \right) = - \frac{\partial}{\partial x_j} q_j \label{eqn:enthalpy-les}
\end{align}
It is noted here that the terms $\hat{q}_j$ (Eq.~\ref{eqn:pot-temp-les}) and
$q_j$ (Eq.~\ref{eqn:enthalpy-les}) are not equivalent and must be scaled
appropriately. User can still provide the appropriate initial and boundary
conditions in terms of potential temperature field. Under these assumptions and
conditions, the resulting solution can then be interpreted as the variation of
potential temperature field in the computational domain.

The subgrid-scale kinetic energy ($k$) one-equation turbulence model
(Eq.~\ref{eqn:ksgs-les}), based on Moeng~\cite{Moeng1984}, is used for
LES turbulence closure.

\begin{align}
  \frac{\partial}{\partial t}\left(\bar{\rho} k \right) + \frac{\partial}{\partial x_j} \bar{\rho} k \widetilde{u_j} &= \frac{\partial}{\partial x_j} \left (2 \nu_t \frac{\partial k}{\partial x_j} \right) - 2 \nu_t \widetilde{S_{ij}} \widetilde{S_{ij}} + \frac{g}{\theta_\circ} \tau_{\theta_i} - C_{\epsilon} \frac{k^{3/2}}{l} \label{eqn:ksgs-les}
\end{align}
where
\begin{align*}
  C_\epsilon &= 0.93 & \nu_t &= C_\epsilon l \sqrt{k} \\
  \tau_{\theta i} &= - 2\ \frac{\nu_t}{\mathrm{Pr}_t}\ \frac{\partial \widetilde{\theta}}{\partial x_i} & \mathrm{Pr}_t &= \left( 1 + \frac{2l}{\Delta} \right)^{-1} \\
\end{align*}
The choice of the length scale ($l$) depends on whether the stratification is positive
or negative, based on the recommendation by Moeng~\cite{Moeng1984}, and is given
by
\begin{align*}
  l =
  \begin{cases}
    \Delta = \sqrt{\Delta_x \Delta_y \Delta_z}, & \frac{\partial \widetilde{\theta}}{\partial z} < 0 \\
    0.76\, \sqrt{k}\, \frac{g}{\theta_\circ}\, \left( \frac{\partial \widetilde{\theta}}{\partial z}\right)^{-1/2}, &  \frac{\partial \widetilde{\theta}}{\partial z} > 0 \\
  \end{cases}
\end{align*}

\begin{comment}
  \fxnote{Shreyas write this part.  Currently placeholder equations below.}\\

  The low-Mach number equations:
  \begin{align}
    \frac{\partial \rho} {\partial t} + \frac{\partial \rho u_j}{\partial x_j} & = 0, \\
    \frac{\partial \rho u_i}{\partial t} + \frac{\partial \rho u_j u_i}{\partial x_j}
    + \frac{\partial P}{\partial x_i} & = \frac{\partial \tau_{ij}}{\partial x_j}
                                        + \left( \rho - \rho_{\circ} \right) g_i, \\
    \frac{\partial \rho h}{\partial t} + \frac{\partial \rho u_j h}{\partial x_j} & =
                                                                                    - \frac{\partial q_j}{\partial x_j} + \frac{\partial P_{th}}{\partial t},
  \end{align}
\end{comment}

@@

\subsubsection{Nalu-Wind}

Nalu-Wind~\cite{SpragueAVR2020} is an open-source, wind-focused fork of the Nalu CFD
code~\cite{Domino:2015}. The codebase heavily leverages the open-source Trilinos
suite of numerical libraries~\cite{Heroux:2003}. Nalu-Wind uses an
unstructured-grid, node-centered, finite-volume formulation for solving the incompressible
Navier-Stokes equations. Two spatial discretization approaches are supported
within the codebase -- the edge-based scheme and control-volume finite-element
scheme. The edge-based scheme, used in the present study, is similar to the
classical cell-centered finite volume formulation formulated on a dual volume
surrounding the nodes of the mesh. The governing equations are discretized in
time using a split-operator approach. Implicit backward difference formula
(BDF2) timestepping algorithm with an approximate pressure projection
scheme~\cite{Moen-Domino:2003} is used to advance the equations in time.
Multiple Picard-style outer-loop iterations are employed within each timestep
to minimize the errors introduced from split-operator approximations. Nalu-Wind
comes equipped with a suite of Reynolds-Averaged Navier-Stokes (RANS), LES, and
Detached Eddy Simulation (DES) turbulence models to resolve the disparate
spatial and temporal scales encountered when modeling wind plant flows.

Nalu-Wind heavily utilizes the Sierra Toolkit (STK) library~\cite{Edwards:2010}
for managing the unstructured mesh data structures. The linear systems resulting
from the discretized equations are solved using solvers available within the
Belos and MueLu libraries within Trilinos. The momentum and scalar transport
equations are solved using GMRES iterative solver with symmetric Gauss-Seidel
(SGS) preconditioners. The elliptic pressure-Poisson equation is solved using
GMRES solvers with a smoothed-aggregated algebraic multigrid preconditioner. The
reader is referred to Sprague et al.~\cite{SpragueAVR2020} and
Domino~\cite{Domino:2015} more more details.

While the primary objective of Nalu-Wind within the ExaWind simulation suite is
to support blade-resolved simulations of wind turbines, it comes equipped
with all the physics models necessary for modeling atmospheric boundary flows.
In the present work, the prognostic equation for temperature is solved using the
enthalpy equation (Eq.~\ref{eqn:enthalpy-les}). The prognostic one-equation,
subgrid-scale kinetic energy equation (Eq.~\ref{eqn:ksgs-les}) is used for LES
turbulence closure. Two Picard iterations were employed within each timestep to
minimize splitting errors and recover second-order accuracy in time.
